\documentclass[12pt, a4paper]{article}

% ── Packages ──────────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{caption}
\usepackage{enumitem}
\usepackage[numbers]{natbib}
\usepackage{setspace}
\usepackage{url}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,calc,fit}

% ── Typography ─────────────────────────────────────────────────────────────────
\onehalfspacing
\setlength{\parskip}{0.5em}

% ── Title ─────────────────────────────────────────────────────────────────────
\title{\textbf{MIVA-KNIGHT: A Domain-Adaptive Multimodal Voice Assistant\\[4pt]
       using Hybrid RAG}}
\author{Oluwakayode (Kayode) Soyinka\\[2pt]
        \small Concordia University of Edmonton\\
        \small IT 571 --- Course Project Proposal}
\date{}

% ==============================================================================
\begin{document}
\maketitle
\thispagestyle{empty}

% ── Summary ───────────────────────────────────────────────────────────────────
\noindent\textbf{Summary}

\noindent
This project aims at developing MIVA-KNIGHT, a domain-adaptive, privacy-ready
voice assistant that focuses on fixing the main flaw that AI assistants out
there have where they prioritize knowing it all over accuracy and depth. It aims
at solving the issue where current models rely on general internet noise and
cannot refine the knowledge needed for privacy-critical sectors like knowing the
in-depth knowledge of company bylaws rather than just knowing they exist.

To make this work, the system relies on a Hybrid RAG architecture~\cite{sarmah2024hybridrag}
that gets the best out of two worlds. It uses Knowledge Graphs to map out
structured, multi-hop relationships in the data, while simultaneously using
vector retrieval to understand unstructured text. This combination ensures the
model grounds its answers in verifiable facts rather than statistical guesses.
The design guarantees secure deployment by running the whole pipeline in a
local, offline environment using containerized microservices~\cite{lewis2020rag}.
For this capstone, the focus is on building a functional prototype to
demonstrate these capabilities. This involves developing a user-friendly web
interface for voice interaction and a backend system that visualizes the AI's
reasoning path.

This approach highlights the novelty of MIVA-KNIGHT where the model is used
in a specific domain with the use of structured knowledge, so it is not focused
on general knowledge like general LLMs but more on depth and accuracy. By
offering this transparent reasoning path, the system improves trustworthiness
and domain adaptability compared to existing multi-modal voice assistants that
operate as black boxes~\cite{kadam2022multimodal}. Finally, the system is
validated against a specific domain dataset to measure its accuracy and
response speed compared to traditional search methods.

\newpage

% ── Table 1 ───────────────────────────────────────────────────────────────────
\begin{table}[h!]
\centering
\caption{Comparative Analysis of Related Work in RAG and Voice Assistants}
\label{tab:related_work}
\renewcommand{\arraystretch}{1.35}
\begin{tabularx}{\textwidth}{@{}p{2.4cm} p{0.8cm} p{3.2cm} p{3.2cm} p{3.6cm}@{}}
\toprule
\textbf{Work} & \textbf{Year} & \textbf{Methodology} & \textbf{Advantages} & \textbf{Limitations} \\
\midrule
Sarmah et al.~\cite{sarmah2024hybridrag}
  & 2024
  & HybridRAG: Combines vector retrieval for unstructured text with Knowledge
    Graphs for structured data.
  & ``Best of both worlds'': high accuracy; reduces hallucinations; grounds
    answers in facts.
  & Complex architecture requiring significant computational resources. \\[4pt]

Edge et al.~\cite{edge2024graphrag}
  & 2024
  & GraphRAG: Uses LLMs to build a KG from documents to answer global summary
    questions.
  & Excellent for holistic questions (e.g., ``What are main themes?'');
    superior comprehensiveness.
  & High token cost due to extensive summarization; graph construction is
    slow. \\[4pt]

Pan et al.~\cite{pan2024unifying}
  & 2024
  & Unified LLM+KG: A roadmap of 3 frameworks: KG-enhanced LLMs,
    LLM-augmented KGs, and Synergized models.
  & Enhances interpretability and reasoning; addresses ``black box'' nature of
    standard LLMs.
  & KG construction is difficult and expensive; real-time evolution is
    challenging. \\[4pt]

Kadam~\cite{kadam2022multimodal}
  & 2022
  & Multimodal Interaction: Framework for combining voice, visual, and touch
    inputs/outputs.
  & Increases user trust by ``showing'' visuals alongside speaking; improves
    usability.
  & Technical challenges in synchronization (e.g., lip-syncing). \\[4pt]

Lewis et al.~\cite{lewis2020rag}
  & 2020
  & Original RAG: Foundational retrieval-augmented generation using dense
    vector indices.
  & Proven superiority over standard models; easier to update knowledge than
    retraining.
  & Can hallucinate if retrieved docs are irrelevant; lacks multi-hop
    reasoning. \\
\bottomrule
\end{tabularx}
\end{table}

\newpage

% ==============================================================================
\section{Introduction}

\subsection{Background and Motivation}

Voice assistants have fundamentally changed how we interact with information,
but they have a major flaw: they prioritize sounding like they know everything
over actually being accurate. While modern AI systems confidently answer broad
questions, they are largely trained on general internet noise. This data quickly
becomes outdated and often leads to hallucinations~\cite{lewis2020rag}. When a
model confidently generates a plausible but factually wrong answer, that
isn't just an edge case --- it is a critical system failure.

In privacy-critical fields like healthcare and corporate governance, this lack
of factual grounding makes ``black box'' systems a serious
liability~\cite{pan2024unifying}. They struggle to synthesize the nuanced
knowledge required for specialized tasks. For example, a standard model might
know that company bylaws exist, but it can't cross-reference the specific,
interlinked clauses needed for actual compliance~\cite{johnson2019mimic}.
MIVA-KNIGHT was built to solve this problem. By grounding every response in
verifiable, domain-specific evidence, it shifts the AI paradigm from
\emph{knowing it all} to \emph{knowing it right}~\cite{sarmah2024hybridrag}.

\subsection{Problem Statement}

Traditional Retrieval-Augmented Generation (RAG) systems were created to reduce
hallucinations by anchoring AI responses to external
documents~\cite{lewis2020rag}. However, standard RAG frameworks have a critical
blind spot: they rely almost entirely on dense vector retrieval. This reliance
creates a complexity wall when the system faces three specific real-world
challenges:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{The Multi-Hop Reasoning Gap:} Vector similarity searches
    struggle with queries that require connecting multiple related concepts. For
    example, figuring out contraindicated treatments across different medical
    conditions requires reasoning through entity relationships, not just pulling
    up isolated documents~\cite{sarmah2024hybridrag}.

  \item \textbf{The Modality Silo:} Most RAG setups are strictly text-centric,
    meaning they can't handle the diverse data like images, audio recordings,
    and sensor readings generated in real-world environments like clinics or
    factories~\cite{kadam2022multimodal}. Real user queries are rarely confined
    to just one modality.

  \item \textbf{The Trust and Privacy Deficit:} Existing multi-modal AI models
    usually operate as black boxes dependent on the cloud, making them a
    non-starter for privacy-sensitive institutional
    data~\cite{pan2024unifying}. High-stakes environments simply cannot trust
    systems that lack transparent, audit-able reasoning paths.
\end{itemize}

\subsection{Research Question and Hypothesis}

This project aims to develop MIVA-KNIGHT, a domain-adaptive, privacy-ready
voice assistant powered by MM-HybridRAG --- a multi-modal extension of the
Hybrid Retrieval-Augmented Generation
framework~\cite{sarmah2024hybridrag}.

\noindent\textbf{Research Question:} How does combining structured knowledge
graph traversal with multi-modal dense retrieval impact the accuracy,
hallucination resistance, and multi-hop reasoning capabilities of a
domain-specific assistant?

\noindent\textbf{Hypothesis:} The core hypothesis is that combining structured
Knowledge Graphs (which map out multi-hop relationships) with unstructured
vector retrieval (which handles semantic meaning) will significantly boost
accuracy and reduce hallucinations compared to traditional vector-only methods.
By implementing this hybrid approach across text
(BERT)~\cite{devlin2019bert}, visual (ResNet-50)~\cite{he2016deep}, and audio
(Wav2Vec~2.0)~\cite{baevski2020wav2vec} modalities, the model can ground its
answers in verifiable facts rather than statistical
guesses~\cite{kadam2022multimodal}.

\subsection{Project Objectives}

To demonstrate these capabilities, the primary goal of this capstone is to
build a functional prototype. This involves four main objectives:

\begin{itemize}[leftmargin=1.5em]
  % ── FIX 1: weights corrected from 60/40 to 70/30 ─────────────────────────
  \item \textbf{Core Architecture:} Build a hybrid retrieval engine that
    combines dense vector similarity (weighted at 70\%) with 2-hop knowledge
    graph traversal (weighted at 30\%)~\cite{sarmah2024hybridrag}. This setup
    ensures the system actually reasons across entity relationships instead of
    just fetching isolated text matches.

  \item \textbf{Multi-Modal Integration:} Create a cross-attention fusion
    mechanism~\cite{kadam2022multimodal} that can ingest and unify text, visual
    data (from the COCO dataset~\cite{lin2014microsoft}), and audio (from the
    RAVDESS corpus~\cite{livingstone2018ryerson}) to handle complex,
    heterogeneous queries.

  % ── FIX 2 + 3: ROCO replaces MIMIC-CXR; P@1 target corrected to >60% ─────
  \item \textbf{Domain Adaptation:} Prove the system can transition from
    general visual data to highly specialized medical imaging using the ROCO
    radiology dataset. The goal here is a Precision@1 exceeding 60\% on the
    ROCO medical benchmark, demonstrating viability for privacy-critical
    deployments where accuracy is non-negotiable.

  \item \textbf{Rigorous Benchmarking:} Evaluate the system using a
    multi-category hallucination test suite targeting greater than 90\%
    hedging and rejection rate, and a multi-hop accuracy evaluation on the
    MultiHop-RAG and HotpotQA
    benchmarks~\cite{sarmah2024hybridrag,tang2024multihop} to confirm it is
    deployment-capable.
\end{itemize}

\subsection{Significance of the Study}

MIVA-KNIGHT stands out because of its targeted approach: by running in a
specific domain with structured knowledge, it values deep, accurate, and secure
answers over broad, surface-level trivia. For security, MIVA-KNIGHT adopts a
privacy-ready architecture: the system is developed using cloud infrastructure
for rapid prototyping and fair benchmarking (enabling direct comparison with
SAM-RAG, RT-RAG, SCM-RAG), then containerized for offline deployment
demonstration in Month~7. This approach balances development speed with the
strict offline constraints of healthcare and corporate governance.

Additionally, the system automatically builds its knowledge graphs based on
Named Entity Recognition (NER) co-occurrence, completely bypassing the need for
manual curation or external
ontologies~\cite{pan2024unifying}. This creates a highly scalable blueprint for
domain-adaptive AI in industries such as healthcare and corporate governance.
Finally, by showing users exactly how it arrived at an answer, the system offers
a level of transparency and trust that legacy black-box voice assistants simply
cannot match~\cite{kadam2022multimodal}.

\subsection{Known Limitations}

As with any research, there are some anticipated technical hurdles. Due to
computational limits, training is currently restricted to the COCO val2017
split (25,000 images)~\cite{lin2014microsoft}, which might limit the
generalization of the model. Also, our automated knowledge graph construction
currently uses word frequency instead of the more robust transformer-based Named
Entity Recognition (NER). This is a known gap for graph
quality~\cite{pan2024unifying} and is the likely reason our initial multi-hop
accuracy sits at a baseline of 30\% rather than our ultimate 80\%
target~\cite{sarmah2024hybridrag}.

On the audio side, the Wav2Vec~2.0 processing~\cite{baevski2020wav2vec} is
currently built for batch processing, not real-time streaming.

% ── FIX 3: MIMIC-CXR credentialing reframed as future work, not a scope gap ──
Finally, the medical validation phase uses the freely available ROCO radiology
dataset rather than MIMIC-CXR~\cite{johnson2019mimic}, which requires
multi-week institutional PhysioNet credentialing. MIMIC-CXR integration is
documented as a structured future enhancement path: once credentials are
obtained, only Pipeline~B (the medical ImageProjection head and FAISS index)
would need retraining, leaving all other system components unchanged. This
design decision removes credentialing as a project scope constraint while
preserving a clear upgrade path.

% ==============================================================================
\section{Literature Review}

\subsection{Overview}

The emergence of large language models (LLMs) has transformed information
retrieval, yet a fundamental flaw persists: LLMs hallucinate at rates of
5--12\%, generating plausible but factually incorrect
answers~\cite{zhang2023sirens,ji2023survey}. In privacy-critical regulated
environments governed by HIPAA and GDPR, even moderate hallucination rates
carry civil liability and risk to public health~\cite{tang2024multihop}.

Retrieval-Augmented Generation (RAG) was introduced by Lewis et
al.~\cite{lewis2020rag} to anchor model outputs to retrieved documents,
reducing parametric guesswork. However, standard dense-vector RAG introduces
critical limitations in multi-hop reasoning and modality support that prevent
deployment in specialized domains. The subsections below synthesize the six core
themes identified across the reviewed literature and map each finding directly
to the MIVA-KNIGHT design.

\subsection{Key Themes and Findings}

The table below consolidates the principal findings across six thematic
clusters, with an explicit link to architectural decisions made in MIVA-KNIGHT.

\begin{longtable}{@{}p{2.0cm} p{4.8cm} p{5.0cm}@{}}
\caption{Summary of Literature Themes and Findings}
\label{tab:themes}\\
\toprule
\textbf{Theme} & \textbf{Key Finding} & \textbf{Relevance to MIVA-KNIGHT} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} --- continued}\\
\toprule
\textbf{Theme} & \textbf{Key Finding} & \textbf{Relevance to MIVA-KNIGHT} \\
\midrule
\endhead
\bottomrule
\endfoot

1.~Hallucination
  & LLMs hallucinate at 5--12\% in general
    use~\cite{zhang2023sirens,ji2023survey}. Even low rates are legally
    unacceptable in HIPAA/GDPR contexts~\cite{tang2024multihop}.
  & Motivates tiered confidence guards and the $>$90\% hallucination
    hedging/rejection target. \\[6pt]

% ── FIX 1: Table 2 row 2 — weights corrected from 60/40 to 70/30 ─────────────
2.~RAG Evolution
  & Lewis et al.~\cite{lewis2020rag} set the dense baseline (45.6\% on
    HotpotQA). Sarmah et al.~\cite{sarmah2024hybridrag} showed hybrid RAG
    (dense + KG) reaches 60--65\%.
  & Informs the 70/30 weighted fusion ($\alpha{=}0.7$ dense, $0.3$ graph)
    core architecture. \\[6pt]

3.~KG Construction
  & Manual curation yields 91--95\% precision but is
    slow~\cite{rossiello2023knowledge}. GraphRAG~\cite{edge2024graphrag}
    requires \$2K--\$5K and 450+ GPU-hours per domain.
  & Justifies NER-based automatic KG achieving $>$80\% coverage in under
    10~minutes. \\[6pt]

4.~Multimodal Fusion
  & HM-RAG~\cite{liu2025hmrag} and SAM-RAG~\cite{zhai2024samrag} process
    multimodal documents but not simultaneous live voice + image + text
    streams.
  & Addresses the real-time gap using BERT~\cite{devlin2019bert},
    ResNet-50~\cite{he2016deep}, and Wav2Vec~2.0~\cite{baevski2020wav2vec}. \\[6pt]

5.~Domain Adaptation
  & Pan et al.~\cite{pan2024unifying} show KG-enhanced LLMs improve domain
    reasoning. Legal (LexGLUE~\cite{chalkidis2022lexglue}) benchmarks prove
    feasibility.
  & Validates retraining only the ImageProjection head ($\sim$2.23\% of
    parameters). \\[6pt]

6.~Evaluation
  & Multi-hop accuracy degrades sharply: 45.6\%
    (2-hop~\cite{yang2018hotpotqa}) $\to$ 23.1\%
    (3.2-hop~\cite{trivedi2022musique}) $\to$ $<$20\%
    (4-hop~\cite{tang2024multihop}).
  & Motivates a unified protocol reporting P@1, MRR, and 4-hop accuracy. \\

\end{longtable}

\subsection{Research Gap}

A comparison of 2024--2025 systems --- SAM-RAG~\cite{zhai2024samrag},
RT-RAG~\cite{shi2026rtrag}, HM-RAG~\cite{liu2025hmrag},
GraphRAG~\cite{edge2024graphrag}, and
OmniSearch~\cite{li2024omnisearch} --- reveals a consistent pattern: progress
has been made in individual capabilities, but no existing system simultaneously
combines offline deployability, multi-modal live fusion, and
privacy-compliant knowledge grounding. Five specific gaps were identified:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Cloud Dependency:} SAM-RAG and OmniSearch rely on OpenAI APIs;
    however, Protected Health Information (PHI) cannot traverse outside
    hospital networks (HIPAA \S\,164.312(e)).

  \item \textbf{KG Cost Bottleneck:}
    GraphRAG~\cite{edge2024graphrag} requires extensive GPU-hours, making rapid
    domain transfer prohibitive for institutions with limited compute budgets.

  \item \textbf{Benchmark Fragmentation:} No two reviewed systems use the same
    evaluation protocol, making cross-system comparison unreliable and
    hindering reproducibility.

  % ── FIX 3: MIMIC-CXR → ROCO in the gap description ──────────────────────
  \item \textbf{Absent Medical P@1 Reporting:} No reviewed system reports P@1
    on open-access medical retrieval benchmarks such as the ROCO radiology
    dataset, despite retrieval precision being a critical metric for clinical
    decision support applications.

  \item \textbf{Modality Limitation:} Most `multimodal' systems process static
    archived documents rather than simultaneous live voice + image + text
    inputs, leaving the real-time interaction scenario unaddressed.
\end{itemize}

\subsection{MIVA-KNIGHT Positioning}

MIVA-KNIGHT is positioned to close all five gaps through four deliberate design
decisions:

\begin{enumerate}[leftmargin=1.5em]
  \item \textbf{Containerized Offline Deployment:} The full pipeline runs
    locally using quantized BERT~\cite{devlin2019bert} and an on-device LLM
    (Llama-3.1-8B via Groq), with no external API calls required at inference
    time.

  \item \textbf{Automatic KG Construction:} NER-based entity extraction
    achieves $>$80\% entity coverage in under ten minutes, eliminating the
    GPU-hour cost of GraphRAG~\cite{edge2024graphrag} without sacrificing
    semantic fidelity.

  \item \textbf{True Cross-Modal Fusion:} All modalities --- text, image, and
    audio --- are projected into a shared 512-dimensional embedding space,
    enabling real-time simultaneous retrieval rather than sequential
    document-level processing.

  \item \textbf{Unified Evaluation Protocol:} Every experimental result is
    reported using the same set of metrics (P@1, P@3, MRR, and multi-hop
    accuracy), enabling direct comparison with prior systems. Month~4 results
    already validate this foundation: P@1\,=\,64.6\%, P@3\,=\,76.8\%, and
    MRR\,=\,0.7269 on the ROCO medical benchmark.
\end{enumerate}

% ==============================================================================
\section{Methodology: The MM-HybridRAG Framework}

\textit{Scope note:} This section describes the methodology for the
\textbf{IT~571 capstone research prototype} implemented as reproducible Jupyter
notebooks and supporting \LaTeX{} documentation. It is distinct from external
funding or scholarship proposals: the capstone prioritizes verifiable
experiments on public benchmarks (COCO, ROCO, RAVDESS, CREMA-D, CMU-MOSI), not
multi-industry production rollouts.

MIVA-KNIGHT is realized as a \textbf{Domain-Adaptive Multimodal
Retrieval-Augmented Generation (MM-HybridRAG)} stack. Parametric LLMs remain
prone to hallucination at an estimated 5--12\%~\cite{zhang2023sirens,ji2023survey};
MM-HybridRAG instead conditions generation on \emph{retrieved} evidence and a
lightweight knowledge graph, with multimodal encoders providing aligned features
for routing, emotion context, and retrieval. Operationally, the work is organized
into \textbf{five training pipelines} (A--E) that share design patterns---frozen
backbones, trainable projection heads or fusion modules, one-time embedding
caches, and unified metrics---followed by a \textbf{hybrid retrieval and
generation} stage that constitutes the capstone integration path.

Figure~\ref{fig:methodology} summarizes the end-to-end methodology from
multimodal input to grounded outputs. The numbered phases align with the
executable notebook structure referenced throughout the project artifacts.

\begin{figure}[h!]
  \centering
  \begin{tikzpicture}[
    font=\small,
    box/.style={rectangle, rounded corners=2pt, draw=black!65, align=center,
      minimum width=2.85cm, minimum height=0.85cm, inner sep=3pt},
    wide/.style={box, minimum width=6.4cm},
    arr/.style={-{Stealth[length=2.2mm]}, thick, black!55},
    lab/.style={font=\scriptsize, inner sep=1pt}
  ]
    % Row 1: inputs
    \node[box, fill=blue!8] (in1) {Voice / audio\\(Whisper ASR)};
    \node[box, fill=blue!8, right=0.35cm of in1] (in2) {Text / query};
    \node[box, fill=blue!8, right=0.35cm of in2] (in3) {Image / frame};
    % Row 2: encoders (wide bar under inputs)
    \node[box, fill=orange!12, below=0.55cm of in2, minimum width=8.6cm] (enc)
      {Frozen encoders + 512-D projection heads (BERT, ResNet-50, Wav2Vec)};
    \node[lab, left=0.1cm of enc.west, anchor=east] {\textbf{1--2}};
    % Pipelines band
    \node[wide, fill=green!10, below=0.45cm of enc] (pipe)
      {\textbf{Pipeline training (capstone artifacts)}\\
      A: COCO text--image \quad B: ROCO transfer \quad C: RAVDESS\\
      D: CREMA-D audio--video \quad E: CMU-MOSI fusion};
    \node[lab, left=0.1cm of pipe.west, anchor=east] {\textbf{2}};
    % Row: emotion + router
    \node[box, fill=yellow!15, below=0.45cm of pipe.west, xshift=-0.8cm] (emo)
      {Emotion / sentiment\\context (C--E)};
    \node[box, fill=yellow!15, right=0.4cm of emo] (route)
      {Domain router\\(general vs.\ medical)};
    \node[lab, left=0.1cm of emo.west, anchor=east] {\textbf{3}};
    % Hybrid retrieval
    \node[box, fill=teal!12, below=0.55cm of $(emo)!0.5!(route)$] (hyb)
      {Hybrid retrieval: FAISS dense +\\co-occurrence KG ($s{=}0.7 s_d{+}0.3 s_{KG}$)};
    \node[lab, left=0.1cm of hyb.west, anchor=east] {\textbf{4}};
    % Confidence
    \node[box, fill=red!8, below=0.45cm of hyb] (conf)
      {Tiered confidence guard\\(green / amber / red)};
    \node[lab, left=0.1cm of conf.west, anchor=east] {\textbf{5}};
    % Generation
    \node[box, fill=violet!10, below=0.45cm of conf] (llm)
      {Grounded generation (Llama-3.1-8B)\\top-$k$ evidence + emotion in prompt};
    \node[lab, left=0.1cm of llm.west, anchor=east] {\textbf{6--7}};
    % Outputs
    \node[box, fill=gray!12, below=0.45cm of llm] (out)
      {Outputs: text, TTS (e.g.\ gTTS),\\cited sources / reasoning trace (UI)};
    \node[lab, left=0.1cm of out.west, anchor=east] {\textbf{7}};
    % Arrows
    \draw[arr] (in1.south) -- ++(0,-0.15) -| (enc.north);
    \draw[arr] (in2.south) -- (enc.north);
    \draw[arr] (in3.south) -- ++(0,-0.15) -| (enc.north);
    \draw[arr] (enc.south) -- (pipe.north);
    \draw[arr] (pipe.south) -- ++(0,-0.2) -| (emo.north);
    \draw[arr] (pipe.south) -- ++(0,-0.2) -| (route.north);
    \draw[arr] (emo.south) |- ($(hyb.north west)!0.3!(hyb.north east)$);
    \draw[arr] (route.south) |- ($(hyb.north west)!0.7!(hyb.north east)$);
    \draw[arr] (hyb.south) -- (conf.north);
    \draw[arr] (conf.south) -- (llm.north);
    \draw[arr] (llm.south) -- (out.north);
  \end{tikzpicture}
  \caption{MIVA-KNIGHT capstone methodology: multimodal encoding, five pipeline
    artifacts (A--E), routing and hybrid retrieval, confidence gating, and
    grounded multi-channel output. Phase labels match
    Section~\ref{sec:method-detail}.}
  \label{fig:methodology}
\end{figure}

\subsection{Multimodal Input and Feature Encoding (Phases 1--2)}
\label{sec:method-detail}

The MM-HybridRAG prototype accepts \textbf{text}, \textbf{speech} (decoded to
text), and \textbf{images} (including video frames where applicable), addressing
the Modality Silo gap~\cite{kadam2022multimodal}. Speech is transcribed with
OpenAI Whisper in the notebook environment; typed queries and captions are fed
directly to the text encoder. Images are processed as still frames or
radiology thumbnails depending on the benchmark.

Each stream is embedded with \textbf{frozen} backbone models and a
\textbf{trainable} projection head: BERT-base-uncased~\cite{devlin2019bert} for
text, ResNet-50~\cite{he2016deep} for vision, and Wav2Vec~2.0~\cite{baevski2020wav2vec}
for audio. All projected vectors live in a \textbf{shared 512-dimensional}
space (L2-normalized where specified in the notebooks) so cosine similarity
scores retrieval and cross-modal alignment. Training follows five pipelines:
\textbf{A}---COCO val2017 text--image InfoNCE; \textbf{B}---ROCO surgical
transfer of the image head while freezing text and shared objectives;
\textbf{C}---RAVDESS audio-only emotion; \textbf{D}---CREMA-D audio--visual
emotion with cross-modal contrastive phases; \textbf{E}---CMU-MOSI multimodal
sentiment with a cross-modal Transformer fusion head. Heavy encoders run
once to build caches (\texttt{.pt} files) before head-only or fusion training,
which is the capstone's practical answer to limited GPU time.

\subsection{Emotion Detection and Domain Routing (Phase~3)}

Phase~3 attaches \textbf{non-generative} side heads derived from pipelines C--E.
The audio emotion path uses Wav2Vec features and trained projection weights
with supervision from RAVDESS and CREMA-D; the multimodal sentiment path uses
Pipeline~E outputs on CMU-MOSI. Detected emotion or sentiment summary text is
\textbf{concatenated to the prompt} so the final LLM can modulate tone without
inventing clinical facts.

The \textbf{domain router} distinguishes general-domain retrieval (COCO-style
captions and entities) from medical retrieval (ROCO captions and RadLex-biased
entity sets) using lightweight rules (keywords) and embedding similarity. Only
the \textbf{image projection head}, FAISS index, and graph vocabulary swap at
domain change; BERT, Wav2Vec, and the text projection remain frozen, matching the
``surgical transfer'' objective stated in the literature review.

\subsection{Hybrid Retrieval (Phase~4)}

Hybrid retrieval combines \textbf{dense} nearest-neighbor search over FAISS with
\textbf{graph-based} re-scoring built from entity co-occurrence (spaCy- or
frequency-derived entities in the notebooks, with medical seed terms for ROCO).
Dense retrieval receives a 70\% weight and graph-derived multi-hop support
receives 30\%, following Sarmah et al.~\cite{sarmah2024hybridrag}:
\begin{equation}
  s \;=\; 0.7 \cdot s_{\mathrm{dense}} \;+\; 0.3 \cdot s_{\mathrm{KG}}.
\end{equation}
This dual pathway targets the Multi-Hop Reasoning Gap~\cite{lewis2020rag} while
remaining implementable without proprietary graph databases in the capstone
environment.

\subsection{Tiered Confidence Guard (Phase~5)}

Before any generative call, retrieved evidence is scored against
\textbf{domain-specific} thresholds inspired by regulated settings
\cite{tang2024multihop}. The guard uses a three-tier policy:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Green (Proceed):} Similarity and graph support exceed medical
    (0.55) or general (0.40) cutoffs; full generation is allowed.
  \item \textbf{Amber (Warning):} Intermediate scores return partial passages and
    an explicit uncertainty disclaimer.
  \item \textbf{Red (Reject):} Evidence is insufficient; the system refuses
    generation and avoids a parametric guess.
\end{itemize}

The guard is evaluated together with the custom hallucination probes listed
under Expected Outcomes; it is a capstone \emph{demonstration} of risk-aware RAG,
not a certified compliance control.

\subsection{Grounded Generation and Multi-Channel Output (Phases 6--7)}

The \textbf{Llama-3.1-8B} instruct model generates answers via the Groq API
during development, conditioned on the \textbf{top five} retrieved segments and
the emotion context string. The capstone treats this as a \textbf{controlled
prototype}: API use enables rapid iteration for the course timeline; swapping
to a \textbf{local} quantized Llama for offline demo is documented as the
integration stretch goal (parallel to containerized deployment in the outcomes
section). Prompting forbids fabricating citations beyond retrieved IDs.

Outputs are designed for the capstone demo: \textbf{(1)} spoken playback via a
lightweight TTS (e.g.\ gTTS); \textbf{(2)} on-screen text; \textbf{(3)} a panel
listing retrieved sources and intermediate scores so instructors can audit
retrieval and routing. This directly supports the Trust and Privacy
Deficit motivation~\cite{pan2024unifying} at prototype fidelity.

% ==============================================================================
\section{Expected Outcomes}

This section states \textbf{capstone-level} deliverables: measurable results tied
to the notebooks and thesis artifacts submitted for IT~571, and a clear
distinction between \emph{what must be demonstrated for the course} and
\emph{optional stretch goals} (e.g., full offline deployment). These outcomes
are \textbf{not} identical to the broader narrative in external scholarship
applications, which may describe multi-sector timelines or impact metrics beyond
the scope of a single-term project.

\subsection{Retrieval and Hybrid RAG}

The primary quantitative evidence for MM-HybridRAG will come from
\textbf{held-out validation} on the same benchmarks used in training:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{ROCO (medical transfer):} Target Precision@1 $>$ 60\%,
    Precision@3 $>$ 75\%, and MRR $>$ 0.70, consistent with the Month~4
    ablation baseline (P@1\,=\,64.6\% text-only) already reported in the
    project write-up.
  \item \textbf{COCO (general domain):} Target MRR $>$ 0.70 on val2017
    text--image retrieval to show that the shared embedding space remains
    usable without retraining BERT.
  \item \textbf{Hybrid ablation:} Document a \textbf{minimum} $+$15\% P@1 improvement
    when graph re-ranking is enabled on top of dense-only retrieval, citing
    the same experimental protocol (same splits, same $k$) for both
    conditions~\cite{sarmah2024hybridrag}.
\end{itemize}

\subsection{Domain Adaptation and Latency}

The capstone will \textbf{demonstrate} domain swap (general $\leftrightarrow$
medical) by reloading the FAISS index, image projection checkpoint, and graph
vocabulary while keeping text and audio encoders frozen. The success criterion
is \textbf{end-to-end swap latency under 5 seconds} on the development machine,
documented with timestamps in the integration notebook or demo script. This
validates the ``surgical transfer'' claim without implying production
load-tested performance.

\subsection{Emotion and Sentiment Pathways}

Emotion and sentiment modules are evaluated on their respective public datasets:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{RAVDESS (Pipeline C):} At least \textbf{6 of 8} emotion
    categories correct above random (12.5\%), using the supervised contrastive
    setup described in the standalone pipeline notebook.
  \item \textbf{CREMA-D (Pipeline D):} At least \textbf{3 of 6} emotions
    above the 16.7\% random baseline after the cross-modal InfoNCE and
    refinement phases, consistent with the logged Phase~2 improvements in the
    CREMA-D notebooks.
  \item \textbf{CMU-MOSI (Pipeline E):} Test Mean Absolute Error (MAE)
    \textbf{below 1.0} on the continuous $[-3,+3]$ sentiment scale for the
    strict multimodal cohort (paired \texttt{.wav} and \texttt{.mp4}), showing
    that cross-modal fusion adds value beyond text-only baselines.
\end{itemize}

\subsection{Hallucination Mitigation and Transparency}

For the course prototype, hallucination resistance is evaluated with a
\textbf{documented} multi-category test suite (out-of-scope queries, unanswerable
medical prompts, fabrication probes, boundary cases, and adversarial prompts).
The target is \textbf{$>$90\% combined} hedging, abstention, or correct
grounded answers when evidence is missing or weak, with the tiered guard
recording the final decision. This is an \textbf{experimental} capstone metric,
not a regulatory certification.

\subsection{Integration and Capstone Demo}

The following items define the \textbf{minimum} integration surface for IT~571:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Reproducible runbook:} A single README or notebook sequence that
    reproduces the main retrieval and fusion metrics from cached embeddings
    (no undisclosed manual steps).
  \item \textbf{Streamlit demo (\texttt{app.py}) or equivalent:} A web interface
    that accepts voice or text input, returns a text answer and optional TTS,
    and displays \textbf{retrieved passages with hybrid scores}, domain route,
    confidence tier, and emotion/sentiment context. A collapsible ``How did I
    get this answer?'' panel satisfies the transparency objective from the
    research gap.
  \item \textbf{Automated checks:} A \textbf{38-test} regression suite
    covering unit, retrieval, domain swap, and confidence paths, reported with
    pass/fail in the final submission (target \textbf{100\%} pass for
    capstone credit).
\end{itemize}

\textbf{Stretch goal (non-blocking):} Package the inference path in a
Dockerfile or FastAPI wrapper with a \textbf{locally} quantized Llama to
illustrate offline deployment; the capstone may rely on Groq during development
if the stretch is not completed, with that limitation stated in the
demonstration.

\subsection{Summary of Quantitative Targets}

Table~\ref{tab:targets} summarises measurable success criteria for the
\textbf{IT~571 capstone}. The same table should be copied into the final report
with \textbf{actual} numbers filled in.

\begin{table}[h!]
\centering
\caption{Quantitative Success Targets (Capstone)}
\label{tab:targets}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{@{}p{5.5cm} p{2.8cm} p{5.0cm}@{}}
\toprule
\textbf{Metric} & \textbf{Target} & \textbf{Rationale} \\
\midrule
ROCO Medical P@1            & $>$ 60\%     & Grounded in Month~4 ablation baseline (64.6\% text-only) \\
ROCO Medical MRR            & $>$ 0.70     & Standard information retrieval threshold \\
COCO General MRR            & $>$ 0.70     & Confirms domain-agnostic embedding space \\
KG re-ranking P@1 delta     & $\geq$ $+$15\%  & Validates hybrid over dense-only RAG \\
Domain swap latency         & $<$ 5s       & Central architectural claim \\
RAVDESS emotion accuracy    & $\geq$ 6/8 emotions & Meaningful above 12.5\% random baseline \\
CREMA-D emotion accuracy    & $\geq$ 3/6 emotions & $>$2.5$\times$ above 16.7\% random baseline \\
Pipeline E (CMU-MOSI) MAE   & $<$ 1.0      & Beat published EF-LSTM baseline \\
Hallucination mitigation    & $>$ 90\%     & Multi-category custom test suite \\
Integration test pass rate  & 100\% (38/38)& Full system reliability \\
\bottomrule
\end{tabularx}
\end{table}

% ==============================================================================
\bibliographystyle{IEEEtran}
\bibliography{Oluwakayode/Ref_Oluwakayode}

\end{document}